\bf Appendix: Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-01-17 appendix.qmd
Updated the PBMC3k section

Following the process in @xia2023 again using the Human Peripheral Blood Mononuclear Cells (PBMC) data. Note that @xia2023 uses a different subset from @sec-pbmc, which contains $31,021$ cells including cell type labels, and the gene expression levels were in the unit of log-transformed UMI count per $10,000$. They focused on three sequencing methods (inDrops, DropSeq, and SeqWell) and four common cell types Cytotoxic T cell, CD4+T cell, CD14+ Monocyte, and B cell.

@@

```{r}
#| label: fig-pbmc-mse-umap
#| fig-pos: H
#| out-height: 50%
#| fig-cap: "Assessing which of the 2 UMAP layouts with different hyperparameter settings (n_neighbors: $30$, min_dist: $0.3$ (red), n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents UMAP layout (a, b). Of the two, layout b is optimal across all binwidths making it the best choice."

>>>>

Following the process in @xia2023 again using the Human Peripheral Blood Mononuclear Cells (PBMC) data. Note that @xia2023 uses a different subset from the PBMC dataset used by @chen2023, which contains $31,021$ cells including cell type labels, and the gene expression levels were in the unit of log-transformed UMI count per $10,000$. They focused on three sequencing methods (inDrops, DropSeq, and SeqWell) and four common cell types Cytotoxic T cell, CD4+T cell, CD14+ Monocyte, and B cell.

@@

```{r}
#| label: fig-pbmc-mse-umap
#| fig-pos: H
#| out-height: 53%
#| fig-cap: "Assessing which of the 2 UMAP layouts with different hyper-parameter settings (n_neighbors: $30$, min_dist: $0.3$ (red); n_neighbors: $80$, min_dist: $0.5$ (blue)) on the PBMC data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the top plot and in the scatterplots represents UMAP layout (a, b). Of the two, layout b is optimal across all binwidths making it the best choice."

### 2025-01-17 appendix.tex
Updated PBMC3k data figure link
Updated the PBMC3k section

Following the process in \citet{xia2023} again using the Human
Peripheral Blood Mononuclear Cells (PBMC) data. Note that
\citet{xia2023} uses a different subset from \textbf{?@sec-pbmc}, which
contains \(31,021\) cells including cell type labels, and the gene
expression levels were in the unit of log-transformed UMI count per
\(10,000\). They focused on three sequencing methods (inDrops, DropSeq,
and SeqWell) and four common cell types Cytotoxic T cell, CD4+T cell,
CD14+ Monocyte, and B cell.

@@

\textbf{?@fig-model-pbmc-author-proj-scdeed} shows layout from the
original published article along with that suggested by scDEED to be
more accurate. The log MSE vs binwidth plot illustrates that our
approach would suggest that scDEED is correct here, that layout b is
more accurately reflecting the cluster structure in the data.

@@

\includegraphics[width=1\textwidth,height=0.5\textheight]{appendix_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}

@@

\caption{\label{fig-pbmc-mse-umap}Assessing which of the 2 UMAP layouts
with different hyperparameter settings (n\_neighbors: \(30\), min\_dist:
\(0.3\) (red), n\_neighbors: \(80\), min\_dist: \(0.5\) (blue)) on the
PBMC data is the better representation using MSE for varying binwidth
(\(a_1\)). Colour used for the lines and points in the left plot and in
the scatterplots represents UMAP layout (a, b). Of the two, layout b is
optimal across all binwidths making it the best choice.}

>>>>

Following the process in \citet{xia2023} again using the Human
Peripheral Blood Mononuclear Cells (PBMC) data. Note that
\citet{xia2023} uses a different subset from the PBMC dataset used by
\citet{chen2023}, which contains \(31,021\) cells including cell type
labels, and the gene expression levels were in the unit of
log-transformed UMI count per \(10,000\). They focused on three
sequencing methods (inDrops, DropSeq, and SeqWell) and four common cell
types Cytotoxic T cell, CD4+T cell, CD14+ Monocyte, and B cell.

@@

Figure~\ref{fig-pbmc-mse-umap} shows layout from the original published
article along with that suggested by scDEED to be more accurate. The log
MSE vs binwidth plot illustrates that our approach would suggest that
scDEED is correct here, that layout b is more accurately reflecting the
cluster structure in the data.

@@

\includegraphics[width=1\textwidth,height=0.53\textheight]{appendix_files/figure-pdf/fig-pbmc-mse-umap-1.pdf}

@@

\caption{\label{fig-pbmc-mse-umap}Assessing which of the 2 UMAP layouts
with different hyper-parameter settings (n\_neighbors: \(30\),
min\_dist: \(0.3\) (red); n\_neighbors: \(80\), min\_dist: \(0.5\)
(blue)) on the PBMC data is the better representation using MSE for
varying binwidth (\(a_1\)). Colour used for the lines and points in the
top plot and in the scatterplots represents UMAP layout (a, b). Of the
two, layout b is optimal across all binwidths making it the best
choice.}
